Structuring College Access: The Market Segment Model and College Board Geomarkets

### 2025-07-09 text/manuscript-full-KS.qmd
revisions + copy edits

In 1983, Robert Zemsky and Penny Odell authored *The Structure of College Choice*. College Board underwrote the project -- providing data and funding -- and published the book as part of their nascent efforts "to help colleges estimate their enrollment potential, efforts which had faltered largely because the smallest geographic unit used in these analyses was the state" [@RN4982, p. x]. Conceptualized as an effort to to "capture and quantify" (p. 11) the knowledge of admissions officers, @RN4982 developed the Market Segment Model by analyzing the SAT score-sending behavior of high school seniors in 1980. The thesis of the Market Segment Model is that student demand for your institution is a function of social class and geography. Therefore, colleges should recruit from territories that contain large populations of your target social class. Geomarkets are geographic borders meant to define local recruiting markets. Figure X panel A shows Geomarkets in the Chicago-land area. As a planning tool for colleges, the Market Segment Model predicts how demand for a particular college varies by Geomarket, based on characteristics of households in the Geomarket.

@@

Geomarkets have been incorporated into the supply-side structure of college access in three ways. Anecdotally, Geomarkets became an organizing principle for how college admissions offices allocate admissions recruiters to territories. For example, Figure X panel B shows that Chicago Geomarkets define recruiting territories for University of Chicago admissions recruiters. Second, Geomarkets are the basis for the College Board Enrollment Planning Service (EPS). EPS software recommends which Geomarkets a college should recruit from and which schools/communities they should prioritize within targeted Geomarkets.

@@

Third, Geomarkets were incorporated into College Board's student list product, Student Search Service. Student lists contain the contact information of prospective students and have been the primary source of "lead generation" in U.S. higher education since 1972, when College Board began selling names [@belkin2019-studata; @list_biz; @RN4728]. <!-- @RN4895 reported that 87% of private and 86% of public four-year institutions purchase student lists. --> Colleges control which prospect profiles they purchase by selecting search filters, such as SAT score, AP score, state, etc. Geomarket filters enable colleges to include or exclude prospects from particular Geomarkets.

@@

Our analyses address the two research questions, which speak to how Geomarkets are utilized within EPS software (RQ1) and within the Student Search Service student list product (RQ2). First, what is the socioeconomic and racial variation between Geomarkets in metropolitan areas and how does this variation change over time? We address this question by spatially joining Geomarket shapefiles to Census data about socioeconomic and racial characteristics. Second, how does the socioeconomic and racial composition of included versus excluded prospects vary when student list purchases filter on particular Geomarkets? This research question is motivated by @RN4982, which recommend that selective colleges target affluent Geomarkets, and by prior research on enrollment management behavior, which finds that selective colleges focus recruiting visits on high schools in affluent, predominantly white communities [@RN3519; @RN4758]. We analyze this research question using data from actual student list purchases, which we collected using public records requests. We analyze student list purchases that include all Geomarkets in a metropolitan area in order to investigate which prospects would be included or excluded had the student list purchase filtered on particular Geomarkets.

The following section introduces background and scholarship about enrollment management and introduces salient concepts from scholarship on quantification. Second, we introduce the case, describing how @RN4982 developed Geomarkets and the Market Segment Model and how these concepts were incorporated into College Board recruiting products. Third, we describe data and methods. Fourth, we present results. [We find that\ldots ONE OR TWO SENTENCES]. Finally, we discuss implications for scholarship. Third-party vendors have structured college access for decades and their influence is growing. The contemporary enrollment management industry is characterized by private equity owned firms selling software-as-service platforms that utilize the same logic as @RN4982, for example EAB's "pipeline analytics" machine learning product [vista_eab_ml]. We propose two fruitful approaches to studying the role of third-party vendors in enrollment management. First, case studies -- like this manuscript and @RN5039 -- utilize primary data collection to conduct granular analyses of particular products or vendors. Second, a growing number of subscription databases (e.g., Orbis, Pitchbook, LSEG) serve the information needs of the investment community by providing data on private firms and acquisitions or investments in private firms. These data sources may provide the basis for a scalable "knowledge infrastructure" [@RN5025] that investigates the role of third-party vendors in education.

@@

While the contemporary EM industry is characterized by software-as-service products sold to colleges by private equity backed firms [@list_biz], College Board products played a pivotal role in transforming recruiting from an in-house process to a process structured by third-party products. Every year, colleges must figure out which prospects should receive promotional material and which high schools to visit. The College Board Student Search Service -- copied by ACT's Educational Opportunity Service -- became the ubiquitous means of converting prospects into leads that could be served targeted recruiting interventions. The College Board Enrollment Planning Service (EPS), launched in 1984, is an early software-as-service product that provides recommendations about which Geomarkets colleges should recruit from and which high schools they should visit within targeted Geomarkets. Both products are applications of quantification that sort students into colleges. Student Search Service includes or excludes prospects from a student list purchases by colleges based on the value of selected filters like Geomarket while EPS encourages colleges to target Geomarkets and high schools based on the characteristics of households. The multi-disciplinary literature on quantification helps conceptualize how third-party products utilize Geomarkets and the potential consequences of this utilization on the inequitable sorting of students into colleges.

@@

@RN5007 [p. 402] define quantification as "the production and communication of numbers." <!-- The effects of quantification are described by the three themes of reactivity (people and orginizations change behavior in response to salient quantitative measures), discipline (quantification disciplines actors to react in particular ways), and authority (quantification often weakens discretion of local decision-makers).--> The effects of quantification are described by the concepts reactivity, discipline, and authority [@RN5007]. Sociological studies on quantification have predominantly explored how numbers are produced, used, and interpreted across institutional contexts such as administration, democratic rule, economics, personal life and education, with a particular emphasis on rankings [@RN5008; @RN5003; @RN5005]. In the field of critical data studies, quantification is central to scholarship that explores how computation and algorithms structure social opportunities [e.g., @RN4772; @RN4849]. Our study is grounded in the systematic investigation of "interactions between different quantification regimes" across sociology and critical data studies, as called for by @RN5008 \[p. 239\], to explore the role of third-party servicers and products in the sorting of students into colleges. In particular, the discussions of correlation and homophily by @RN4975 show how social reproduction is encoded in algorithms that direct future resources. We explore how correlation and homophily were incorporated into the construction and application of Geomarkets within the Market Segment Model [@RN4982].

@@

@RN4975 problematizes the idea that homophily is a naturally occurring phenomenon. Because network science models -- like correlational models above -- typically do not observe how historical structures affect relationships, they "erase historical contingencies, institutional discrimination, and economic realities" [@RN4975, p. 95] that cause behavior consistent with homophily. Second, in commercial social networks, homophily is more than an assumption; rather, it is programmed into algorithms that create connections between users. Thus, "networks create and spawn the reality they imagine; they become self-fulfilling prophecies" [@Chun2018Queerying, p. 66]. We observe similar phenomena in our case. Based on a network analysis of 1980 SAT score-sending, @RN4982 concludes that like-colleges compete for like-students (i.e., students of similar social origin). In turn, @RN4982 reasons that colleges should target the Geomarkets and high schools targeted by peer colleges. This logic is subsequently programmed into EPS software that recommends which Geomarkets and high schools to recruit from. Thus, homophily observed on the demand side -- which is a function of historic structural inequality –- is programmed into the supply-side.

@@

We make targeted contributions to scholarship on quantification in the sociology of education. As @RN3519 demonstrates in the chapter *Numbers*, EM is fundamentally concerned with quantification. However, sociological analyses of EM focus on the behavior of colleges and their agents, creating the sense that inequalities created by EM are a function of individual colleges reacting to macro structures in their environment such as rankings and financial pressures [e.g., @RN5003; @RN3519]. To our knowledge, the sociology of education has not considered producer-facing products that quantify students on behalf of producers. Market research and investigative journalism indicates that the vast majority of colleges purchase third-party EM products to identify prospects and to decide which schools and communities to target [@RN5027]. Recruiting products grounded in the logic of predictive analytics take a snapshot of student demand -- without considering historical structures that produce inequality -- and then recommend that colleges divert recruiting resources to localities with strong student demand. Based on a snapshot of existing social stratification, market research matches vertically categorized consumers to vertically categorized producers, thereby amplifying the effect of initial stratification on subsequent stratification. Third-party EM products have been sold since the 1980s and have become more abundant with the ubiquity of software-as-service platforms, but have remained invisible to sociology.

```{=html}
<!--  We make third-party EM products the object of empirical analysis, though we cannot claim to assess how the population of colleges utilize these products.
# REMOVED FROM PARAGRAPH ABOVE: First, in the case of college rankings, extant research shows how consumer-facing quantification disciplines the application/enrollment behavior of prospective students, the hiring behavior of employers, and the behavior of colleges that are disciplined to pursue customers with characteristics valued by rankings [@RN5003; @RN2356; @RN2375].  -->
```
This manuscript analyzes the Market Segment Model [@RN4982], which categorized high school students into vertical market segments -- local, in-state, regional, national -- and simultaneously created local Geomarkets that could be evaluated based on their composition of student market segments. The Market Segment Model and Geomarkets became the basis for the College Board *Enrollment Planning Service* (EPS), which advised colleges which Geomarkets to target. Later, Geomarkets were incorporated into the College Board student list product named Student Search Service. Unlike the analysis of UK school league tables by @RN5005, we cannot show the effect of quantification on social reproduction. However, by demonstrating which students are included versus excluded in student list purchases that filter on particular Geomarkets, our analyses provide novel insight into the underlying mechanism by which quantification can reproduce historical class-based and race-based inequality in the sorting of students into colleges.

@@

The primary data source for our second research question is student lists purchased from College Board for which we received both (a) the order summary data and (b) the de-identified prospect-level student list data. This sample includes a total of 414 orders associated with 2,549,085 prospects made by 14 \[KS CHECK THIS\] public universities located in California, Illinois, Texas, and Arizona. Although these lists were purchased by individual universities, the set of prospects included in each list is a function of the search criteria specified for that student list purchase. The resulting student lists could hypothetically extend to any university placing student lists orders based on the same filter criteria. In other words, any observed patterns in the findings below would remain constant for any university selecting the same filters.[^18] Therefore, we utilize these data not to analyze the behavior of individual universities, but to identify which prospective students are included when a particular set of search criteria are selected.

@@

The variables of interest are Geomarket and prospect-level characteristics, which are derived from the pretest questionnaire administered to College Board test-takers that are included in the de-identified student lists we received (e.g., HS code, HS GPA range, intended major). College Board asks students separate questions about ethnicity (Cuban, Mexican, Puerto Rican, other Hispanic, non-Hispanic, ethnicity non-response) and race (American Indian or Alaska Native, Asian, Black, Native Hawaiian or other Pacific Islander, White, race non-response), allowing students to check as many boxes as they want (College Board, 2016) \[CITE\]. From these responses, a College Board "derived aggregate race/ethnicity" was re-created based on the U.S. Department of Education reporting guidelines that includes the following categories: no response; American Indian/Alaska Native; Asian; Black; Hispanic/Latino; Native Hawaiian or Other Pacific Islander; White; other; two or more races, non-Hispanic (College Board, 2016) \[CITE\].[^20] Information about student socioeconomic status is limited. However, we create a measure of whether the prospect is first-generation college student based on their parents' level of education that was self-reported on the College Board questionnaire \[KS double check this\].

@@

Case selection was informed by purposeful sampling to identify "information-rich" cases that can provide knowledge about variation across Geomarkets and over time \[@RN4011; @RN4549\] . We considered metropolitan areas as information rich cases across three broad dimensions: 1) urban model structures, 2) geographical variation, 3) variation across racial/ethnic and socioeconomic characteristics, and 4) availability of student list orders and de-identified list data. Based on @RN4982, we identified classic urban model structures that reflected "major metropolitan area\[s\] \ldots composed of several markets, usually corresponding to the inner city, a first ring of suburbs, and an outer ring of suburbs" (p. 11-12), such as Boston, Chicago, Detroit, Baltimore, Denver, Houston, Miami, and Philadelphia. To compare and contrast with classic urban model structures, we also identified alternative urban models such as bi-modal or metroplex models that contain two centralized cores surrounded by the more classing suburban ring (e.g., Dallas-Fort Worth and San Francisco) and multi-nuclei models where urban growth develops around multiple, sometimes uncentralized and independent, metropolitan areas (e.g., Los Angeles and New York City). The third consideration addresses geographical variations in demographic populations by intentionally selecting metropolitan cases across different U.S. regions (West, Midwest, Northeast, South) to ensure resident and prospect pools mirror broader populations. Related to all previous selection dimensions, we are also interested in variation of racial and socioeconomic populations between Geomarkets within a metropolitan areas (RQ1) and how such variation shapes which prospective students are targeted by student list purchases (RQ2). We therefore considered whether or not metropolitan areas provided sufficient and varied racial and socioeconomic diversity for our research purposes.

@@

@@

>>>>

In 1983, Robert Zemsky and Penny Odell authored *The Structure of College Choice*. College Board underwrote the project -- providing data and funding -- and published the book as part of their nascent efforts "to help colleges estimate their enrollment potential, efforts which had faltered largely because the smallest geographic unit used in these analyses was the state" [@RN4982, p. x]. Conceptualized as an effort to to "capture and quantify" (p. 11) the knowledge of admissions officers, @RN4982 developed the Market Segment Model by analyzing the SAT score-sending behavior of high school seniors in 1980. The thesis of the Market Segment Model is that student demand for your institution is a function of social class and geography. Therefore, colleges should recruit from territories that contain large populations of your target social class. Geomarkets are geographic borders meant to define local recruiting markets. The left panel of @fig-chicago-recruiting shows Geomarkets in the Chicago-land area. As a planning tool for colleges, the Market Segment Model predicts how demand for a particular college varies by Geomarket, based on characteristics of households in the Geomarket.

@@

Geomarkets have been incorporated into the supply-side structure of college access in three ways. Anecdotally, Geomarkets became an organizing principle for how college admissions offices allocate admissions recruiters to territories. For example, @fig-chicago-recruiting shows that Chicago Geomarkets define recruiting territories for University of Chicago admissions recruiters. Second, Geomarkets are the basis for the College Board Enrollment Planning Service (EPS). EPS software recommends which Geomarkets a college should recruit from and which schools/communities they should prioritize within targeted Geomarkets.

@@

Third, Geomarkets were incorporated into College Board's student list product, Student Search Service. Student lists contain the contact information of prospective students and have been the primary source of "lead generation" in U.S. higher education since 1972, when College Board began selling names [@belkin2019-studata; @list_biz; @RN4728]. <!-- @RN4895 reported that 87% of private and 86% of public four-year institutions purchase student lists. --> Colleges control which prospect profiles they purchase by selecting search filters, such as SAT score, AP score, and state. Geomarket filters enable colleges to include or exclude prospects from particular Geomarkets.

@@

Our analyses address two research questions, which speak to how Geomarkets are utilized within EPS software (RQ1) and within the Student Search Service student list product (RQ2). First, what is the socioeconomic and racial variation between Geomarkets in metropolitan areas and how does this variation change over time? We address this question by spatially joining Geomarket shapefiles to Census data about socioeconomic and racial characteristics. Second, how does the socioeconomic and racial composition of included versus excluded prospects vary when student list purchases filter on particular Geomarkets? This research question is motivated by @RN4982, which recommend that selective colleges target affluent Geomarkets, and by prior research on enrollment management behavior, which finds that selective colleges focus recruiting visits on high schools in affluent, predominantly White communities [@RN3519; @RN4758]. We analyze this research question using data from actual student list purchases, which we collected using public records requests. We analyze student list purchases that include all Geomarkets in a metropolitan area in order to investigate which prospects would be included or excluded had the student list purchase filtered on particular Geomarkets.

The following section introduces background and scholarship about enrollment management and introduces salient concepts from scholarship on quantification. Second, we introduce the case, describing how @RN4982 developed Geomarkets and the Market Segment Model and how these concepts were incorporated into College Board recruiting products. Third, we describe data and methods. Fourth, we present results. Findings for RQ1 reveal that Geomarkets mirror deep-rooted patterns of racial and class segregation, with Black and Latinx communities historically and often still concentrated in the poorest Geomarkets, though these dynamics vary by metropolitan area. Findings for RQ2 indicate student list purchases that exclude working- and middle-class Geomarkets are likely to systematically overlook high-achieving Black and Hispanic prospective students, especially those who are first-generation college students, thereby highlighting how third-party recruitment tools serve as mechanisms of social reproduction by sorting students in ways that reinforce historic inequalities in educational opportunity.

Finally, we discuss implications for scholarship. Third-party vendors have structured college access for decades and their influence is growing. The contemporary enrollment management industry is characterized by private equity owned firms selling software-as-service platforms that utilize the same logic as @RN4982, for example EAB's "pipeline analytics" machine learning product [@vista_eab_ml]. We propose two fruitful approaches to studying the role of third-party vendors in enrollment management. First, case studies -- like this manuscript and @RN5039 -- utilize primary data collection to conduct granular analyses of particular products or vendors. Second, a growing number of subscription databases (e.g., Orbis, Pitchbook, LSEG) serve the information needs of the investment community by providing data on private firms and acquisitions or investments in private firms. These data sources may provide the basis for a scalable "knowledge infrastructure" [@RN5025] that investigates the role of third-party vendors in education.

@@

While the contemporary EM industry is characterized by software-as-service products sold to colleges by private equity backed firms [@list_biz], College Board products played a pivotal role in transforming recruiting from an in-house process to a process structured by third-party products. Every year, colleges must figure out which prospects should receive promotional material and which high schools to visit. The College Board Student Search Service -- copied by ACT's Educational Opportunity Service -- became the ubiquitous means of converting prospects into leads that could be served targeted recruiting interventions. The College Board Enrollment Planning Service (EPS), launched in 1984, is an early software-as-service product that provides recommendations about which Geomarkets colleges should recruit from and which high schools they should visit within targeted Geomarkets. Both products are applications of quantification that sort students into colleges. Student Search Service includes or excludes prospects from student list purchases by colleges based on the value of selected filters like Geomarket while EPS encourages colleges to target Geomarkets and high schools based on the characteristics of households. The multi-disciplinary literature on quantification helps conceptualize how third-party products utilize Geomarkets and the potential consequences of this utilization on the inequitable sorting of students into colleges.

@@

@RN5007 [p. 402] define quantification as "the production and communication of numbers." <!-- The effects of quantification are described by the three themes of reactivity (people and orginizations change behavior in response to salient quantitative measures), discipline (quantification disciplines actors to react in particular ways), and authority (quantification often weakens discretion of local decision-makers). The effects of quantification are described by the concepts reactivity, discipline, and authority [@RN5007]. --> Sociological studies on quantification have predominantly explored how numbers are produced, used, and interpreted across institutional contexts such as administration, democratic rule, economics, personal life, and education, with a particular emphasis on rankings [@RN5008; @RN5003; @RN5005]. In the field of critical data studies, quantification is central to scholarship that explores how computation and algorithms structure social opportunities [e.g., @RN4772; @RN4849]. Our study is grounded in the systematic investigation of "interactions between different quantification regimes" across sociology and critical data studies, as called for by @RN5008 \[p. 239\], to explore the role of third-party servicers and products in the sorting of students into colleges. In particular, the discussions of correlation and homophily by @RN4975 show how social reproduction is encoded in algorithms that direct future resources. We explore how correlation and homophily were incorporated into the construction and application of Geomarkets within the Market Segment Model [@RN4982].

@@

@RN4975 problematizes the idea that homophily is a naturally occurring phenomenon. Because network science models -- like correlational models above -- typically do not observe how historical structures affect relationships, they "erase historical contingencies, institutional discrimination, and economic realities" [@RN4975, p. 95] that cause behavior consistent with homophily. Second, in commercial social networks, homophily is more than an assumption; rather, it is programmed into algorithms that create connections between users. Thus, "networks create and spawn the reality they imagine; they become self-fulfilling prophecies" [@Chun2018Queerying, p. 66]. We observe similar phenomena in our case. Based on a network analysis of 1980 SAT score-sending, @RN4982 concludes that like-colleges compete for like-students (i.e., students of similar social origin). In turn, @RN4982 reasons that colleges should target the Geomarkets and high schools targeted by peer colleges. This logic is subsequently programmed into EPS software that recommends which Geomarkets and high schools to recruit from. Thus, homophily observed on the demand side of college choice-- which is a function of historic structural inequality –- is programmed into the supply-side.

@@

We make targeted contributions to scholarship on quantification in the sociology of education. As @RN3519 demonstrates in the chapter *Numbers*, EM is fundamentally concerned with quantification. However, sociological analyses of EM focus on the behavior of colleges and their agents, creating the sense that inequalities created by EM are a function of individual colleges reacting to macro structures in their environment such as rankings and financial pressures [e.g., @RN5003; @RN3519]. To our knowledge, the sociology of education has not considered producer-facing products that quantify students on behalf of colleges. Market research and investigative journalism indicates that the vast majority of colleges purchase third-party EM products to identify prospects and to decide which schools and communities to target [@RN5027]. Recruiting products grounded in the logic of predictive analytics take a snapshot of student demand -- without considering historical structures that produce inequality -- and then recommend that colleges divert recruiting resources to localities with strong student demand. Based on a snapshot of existing social stratification, market research matches vertically categorized consumers to vertically categorized producers, thereby amplifying the effect of initial stratification on subsequent stratification. Third-party EM products have been sold since the 1980s and have become more abundant with the ubiquity of software-as-service platforms, but have remained invisible to sociology.

```{=html}
<!--  We make third-party EM products the object of empirical analysis, though we cannot claim to assess how the population of colleges utilize these products.
# REMOVED FROM PARAGRAPH ABOVE: First, in the case of college rankings, extant research shows how consumer-facing quantification disciplines the application/enrollment behavior of prospective students, the hiring behavior of employers, and the behavior of colleges that are disciplined to pursue customers with characteristics valued by rankings [@RN5003; @RN2356; @RN2375].  -->
```
This manuscript analyzes the Market Segment Model [@RN4982], which categorized high school students into vertical market segments -- local, in-state, regional, national -- and simultaneously created local Geomarkets that could be evaluated based on their composition of student market segments. The Market Segment Model and Geomarkets became the basis for the College Board *Enrollment Planning Service* (EPS), which advises colleges which Geomarkets to target. Later, Geomarkets were incorporated into the College Board student list product named Student Search Service. Unlike the analysis of UK school league tables by @RN5005, we cannot show the effect of quantification on social reproduction. However, by demonstrating which students are included versus excluded in student list purchases that filter on particular Geomarkets, our analyses provide novel insight into the underlying mechanism by which quantification can reproduce historical class-based and race-based inequality in the sorting of students into colleges.

@@

The primary data source for our second research question is student lists purchased from College Board for which we received both (a) the order summary data and (b) the de-identified prospect-level student list data. This sample includes a total of 414 orders associated with 2,549,085 prospects made by 14 public universities located in California, Illinois, Texas, and Arizona. Although these lists were purchased by individual universities, the set of prospects included in each list is a function of the search criteria specified for that student list purchase. The resulting student lists could hypothetically extend to any university placing student lists orders based on the same filter criteria. In other words, any observed patterns in the findings below would remain constant for any university selecting the same filters.[^18] Therefore, we utilize these data not to analyze the behavior of individual universities, but to identify which prospective students are included when a particular set of search criteria are selected.

@@

The variables of interest are Geomarket and prospect-level characteristics, which are derived from the pretest questionnaire administered to College Board test-takers that are included in the de-identified student lists we received (e.g., HS code, HS GPA range, intended major). College Board asks students separate questions about ethnicity (Cuban, Mexican, Puerto Rican, other Hispanic, non-Hispanic, ethnicity non-response) and race (American Indian or Alaska Native, Asian, Black, Native Hawaiian or other Pacific Islander, White, race non-response), allowing students to check as many boxes as they want [@RN0010]. From these responses, a College Board "derived aggregate race/ethnicity" was re-created based on the U.S. Department of Education reporting guidelines that includes the following categories: no response; American Indian/Alaska Native; Asian; Black; Hispanic/Latino; Native Hawaiian or Other Pacific Islander; White; other; two or more races, non-Hispanic [@RN0010].[^20] Information about student socioeconomic status is limited. However, we create a measure of whether the prospect is first-generation college student based on their parents' level of education that was self-reported on the College Board questionnaire.

@@

Case selection was informed by purposeful sampling to identify "information-rich" cases that can provide knowledge about variation across Geomarkets and over time [@RN4011; @RN4549] . We considered metropolitan areas as information rich cases across three broad dimensions: 1) urban model structures, 2) geographical variation, 3) variation across racial/ethnic and socioeconomic characteristics, and 4) availability of student list orders and de-identified list data. Based on @RN4982, we identified classic urban model structures that reflected "major metropolitan area\[s\] \ldots composed of several markets, usually corresponding to the inner city, a first ring of suburbs, and an outer ring of suburbs" (p. 11-12), such as Boston, Chicago, Detroit, Baltimore, Denver, Houston, Miami, and Philadelphia. To compare and contrast with classic urban model structures, we also identified alternative urban models such as bi-modal or metroplex models that contain two centralized cores surrounded by the more classing suburban ring (e.g., Dallas-Fort Worth and San Francisco) and multi-nuclei models where urban growth develops around multiple, sometimes uncentralized and independent, metropolitan areas (e.g., Los Angeles and New York City). The third consideration addresses geographical variations in demographic populations by intentionally selecting metropolitan cases across different U.S. regions (West, Midwest, Northeast, South) to ensure resident and prospect pools mirror broader populations. Related to all previous selection dimensions, we are also interested in variation of racial and socioeconomic populations between Geomarkets within a metropolitan areas (RQ1) and how such variation shapes which prospective students are targeted by student list purchases (RQ2). We therefore considered whether or not metropolitan areas provided sufficient and varied racial and socioeconomic diversity for our research purposes.

@@

```{r fig-chicago-recruiting, echo = FALSE, fig.align = 'center', fig.cap = "The College Board Enrollment Planning Service Chicago Geomarkets (left) and University of Chicago Recruiting Territories (right)", out.width = "85%"}
knitr::include_graphics('./../assets/images/chicago_recruiting.png')
#knitr::include_graphics('./../../outputs/figures/enrollment_funnel.png')
#![The enrollment funnel](assets/images/enrollment_funnel.png)
```

@@

\newpage

# Appendix

```{r market-segment-characteristics, echo = F}
#| label: tbl-market-segment-characteristics
#| tbl-cap: 'Simplified market segment profile, Connecticut Market 3: Fairfield County'

market_segment <- data.frame(
  activity = c('Total test takers','Avg SAT (verbal + math)','Avg # scores sent per test taker','Percent in top 20% of HS class','Percent aspiring to more than BA','Percent family income more than $35,000','Percent both parents with BA'),
  local = c(550,770,2.8,27.8,30.6,11.2,9.0),
  in_state = c(1199,850,3.5,26.1,41.5,20.6,16.3),
  regional = c(1664,970,4.8,44.7,54.5,41.9,34.0),
  national = c(3766,980,5.3,45.7,62.2,43.0,37.1)
)

kable(
  x = market_segment,
  col.names = c('Characteristic','Local','In-state','Regional','National'),
  #caption = 'Simplified market segment profile\\\\ Connecticut Market 3: Fairfield County'
)
```

```{r sample-institutional-profile, echo = F}
#| label: tbl-sample-institutional-profile
#| tbl-cap: 'Simplified sample institutional profile for anonymous institution, students from Connecticut Market 3: Fairfield County'

sample_inst <- data.frame(
  characteristic = c('Total number of scores received','Pct of all test-takers in segment','lib_arts_num','lib_arts_share','engineering_num','engineering_share'),
  local = c(1,0.2,0,0.0,0,0.0),
  in_state = c(58,4.8,25,8.4,2,3.0),
  regional= c(69,4.1,35,7.0,5,5.8),
  national = c(109,2.9,61,5.3,3,0.8),
  total = c(237,3.3,121,5.9,10,1.8)
)

kable(
  x = sample_inst,
  col.names = c('','Local','In-state','Regional','National','Total'),
  #caption = 'Simplified sample institutional profile\\\\ Connecticut Market 3: Fairfield County'
)
